---
title: "spatialR_scratch"
author: "Austin Covey"
date: "4/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load packages}
library(sf)
library(data.table)
library(tigris)
library(ggplot2)
library(maps)
library(tmap)
library(USAboundaries)
library(maptools)
library(tidyverse)
library(stringr)
```

```{r, load data}
cpuc_file   = 'better_cpuc_compiled.csv'
#df <- fread(cpuc_file) #, colClasses = c('character', df$county_fips))

df <- fread(cpuc_file, header = T, colClasses = c('character', rep(NA, 11)))
# best practice is to classify column as a character as you read the data in.

#rep(NA,11)= NA, NA, NA...... 11 times


# We lack leading zeroes (best practice to have them).
```

Now we gotta get the county shapefiles from tigris.
However the code block below gives an error.

This stack overflow post has the same error:
https://stackoverflow.com/questions/61282572/cant-read-shp-file-in-r

```{r, retrieve county data}
ca_counties = counties(state = 'california')
head(ca_counties)
```

Note: sf package does not provide downloads for shape files, it lets you manipulate geometry.

This tigirs website:

https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2019&layergroup=Counties+%28and+equivalent%29

...is where I got county data from. I am gonna store this data in a subdirectory in our working repo.


Ask Sophia about how to generalize file path names in R for collaboration.

```{r, let's use a shape file from the repo}
shape_file <-"C:\\Users\\acove\\Desktop\\GitHub_repos\\caeecc-energy-efficiency\\notebooks\\r-notebooks\\spatial_r\\mapping_attempt1\\2019_county_shapefiles\\CA_Counties"

original_data<-st_read(shape_file)
```

```{r, let's use a shape file from the repo pt 2}


original_data<-st_read("C:\\Users\\acove\\Desktop\\GitHub_repos\\caeecc-energy-efficiency\\notebooks\\r-notebooks\\spatial_r\\mapping_attempt1\\2019_county_shapefiles\\CA_Counties", "CA_Counties_TIGER2016.shp.xml")
```

```{r}
colnames(original_data)
st_geometry(original_data)
```

```{r, Start Here}

#make a data frame
all_counties<- map_data("county")
# Lot of observations, this stores geometries of county polygons in an inefficient way (there are 3000 counties in US, but way more observations (storing in point, line way))

#visualize
ggplot(data=all_counties)+
  geom_polygon(aes(x = long, y = lat, fill = region, group = group))+ #if you were using a different coord system you would have to change long and lat
  # fill indicates what we want to colored in, in this case we're doing the states (region=state)
  # group is extremely important. There are lots of point values in this data. Group tells R that all of these points belong in a county
  coord_fixed(1.3)+
  guides(fill = FALSE) #if change to fill = any other color, then you would see the outline of all of the counties (county lines are there even when you put FALSE; however we don't see them at this granularity)

```
```{r}
head(all_counties)
# group gets you county
# region gets you state
# don't mess with the order, ever (tells R what way the points go)
# like the maps at macaroni grill
# Let's see what is included in this data sets
```
```{r}
table(all_counties$region)
sort(table(all_counties$subregion), decreasing = TRUE)[1:5] #GIVES US TOP 5 COUNTIES by points; this gives us a frequency count of the counties within the US
```

```{r, subset to CA}
CA_counties<- subset(x = all_counties, subset = region == "california")
# Check subset with: table(CA_counties$region)
```

```{r, check that we can merge counties because you can merge your data sucessfully and it can still be totally wrong}
#the length should be 58 for both since there are 58 counties in total 
length(table(CA_counties$subregion))
length(table(df$County))
```

```{r, make lowercase}
#the county names don't match (our data has capitalized first letters County =/= county), so we have to change it
df_edited<-df
df_edited$County <- str_to_lower(df_edited$County)
head(df_edited) #check to see if the county names did change
```

```{r, merge on count name}
#check dimensions before the merge 
dim(df_edited)
dim(CA_counties)

#merge
merged_data<- left_join(CA_counties, df_edited, by = c("subregion"="County")) #the left_join takes all the unique values from left one and only add the matches from the right (doesn't matter in this case because they WILL match since we checked already)
dim(merged_data) 
```
```{r, visualize}

#Mean Income
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `Mean Income`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("Mean Household Income by County")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#there are ways to change colors and get rid of the axis (read more into ggplot:http://sape.inf.usi.ch/quick-reference/ggplot2/colour)

```
```{r, visualize}

#Population
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `Population`))+ 
  coord_fixed(1.3)+
  ggtitle("County Population")+ 
  theme_classic()+ 
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

```

```{r, visualize}

#Energy Savings
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `TotalFirstYearGrosskWh`))+ 
  coord_fixed(1.3)+
  ggtitle("Energy Savings by County")+ 
  theme_classic()+ 
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

```

```{r, visualize}

#Budget
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `Budget`))+ 
  coord_fixed(1.3)+
  ggtitle("Budget by County")+ 
  theme_classic()+ 
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

```
```{r, visualize}

#IRR
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `IRR`))+ 
  coord_fixed(1.3)+
  ggtitle("Relative Rurality by County")+ 
  theme_classic()+ 
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

```

```{r, visualize}

#Tax Revenue
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `Tax Revenue (2016-2018)`))+ 
  coord_fixed(1.3)+
  ggtitle("Tax Revenue by County")+ 
  theme_classic()+ 
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

```

```{r, visualize}

#CES Score
ggplot(data = merged_data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `CES Score Median`))+ 
  coord_fixed(1.3)+
  ggtitle("CES Score Median by County")+ 
  theme_classic()+ 
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text=element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

```