---
title: "Regressions (Clean)"
author: "Michelle Le"
date: "6/20/2021"
output: html_document
---
##Load data and libraries 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = "~/Desktop/caeecc-energy-efficiency/for github/data")
knitr::opts_knit$set(root.dir = "/Users/ranjitster/Dropbox/cetlab_repos/underserved_public_sector_EE/caeecc-energy-efficiency/for github/data")
```

```{r}
K12 <- read.csv("k12_counties_3.csv")
counties <- read.csv("counties_3.csv")
cities <- read.csv("cities_3.csv")
```

```{r, load packages}
library(readr)
library(data.table)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(psych) #for pair.panels()
library(car) #for the qqPlot function
library(nlme) #for fitting mixed effects models (linear models with random effects)
library(fitdistrplus)
library(stargazer) #to make regression tables, learn more at https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf
library(pscl)
library(splines)

#for spatial purposes 
library(sf)
library(tigris)
library(maps)
library(tmap)
library(USAboundaries)
library(maptools)
library(stringr)
library(viridis)
library(bit64)
library(gridExtra)
library(grid)
library(lattice)
library(GISTools)

```

```{r, cite packages}
citation("readr")
citation("data.table")
citation("dplyr")
citation("ggplot2")
citation("tidyverse")
citation("psych")
citation("car") 
citation("nlme") 
citation("fitdistrplus")
citation("stargazer") 

#for spatial purposes 
citation("sf")
citation("tigris")
citation("maps")
citation("tmap")
citation("USAboundaries")
citation("maptools")
citation("stringr")
citation("viridis")
citation("bit64")
citation("gridExtra")
citation("grid")
citation("lattice")
citation("GISTools")
```

```{r, program summary}

claims_category <- read.csv("cpuc_claims_aggregated_by_category_2017_2019.csv")
claims_program <- read.csv("cpuc_claims_aggregated_by_program_2017_2019.csv")


p_claims_no_programs <- ggplot(claims_category, aes(category, PrgID)) +
  geom_bar(stat = 'identity', width = 0.7) +
  labs(y = "Number of Public Sector \nEnergy Efficiency Programs \n2017-2019", x = NULL) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), legend.title = element_blank())


p_claims_costs <- ggplot(claims_category, aes(category, TotalGrossMeasureCost/10^6)) +
  geom_bar(stat = 'identity', width = 0.7) +
  labs(y = "Public Sector Energy Efficiency \nProgram Budget 2017-2019 \n (million USD)", x = NULL) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), legend.title = element_blank())



g_claims_costs <- ggplotGrob(p_claims_costs + labs(tag = "B"))

g_claims_no_programs <- ggplotGrob(p_claims_no_programs + labs(tag = "A")) 

grob_claims <- arrangeGrob(grobs = list(cbind(g_claims_no_programs, g_claims_costs, size = "first")), ncol = 1)

grid.newpage()
grid.draw(grob_claims)

ggsave(file = paste0("claims_summary", ".png"), grob_claims, width=6, height=4)
```

```{r, Start}
#read in data for K-12 EE programs 
K12 <- read.csv("k12_counties_3.csv")
counties <- read.csv("counties_3.csv")
cities <- read.csv("cities_3.csv")
```

##Important 
Things to look for: Small AIC, Large R^2, p-value should be < 0.05 to be significant 

Every group is treated differently, but there is some similarity in the workflow and how regressions were set up. Chunks in each section follows as so:


2. Perform necessary transformations of variables to use for later (log-transformation, rounding, etc.)
3. Subset dataset to groups we want to analyze (those who participated, those who didn't, quantiles, omitting NAs (city-specific))
4. Identify outliers through the use of Cook's distance (using the 4/n cutoff) and take them out (K12_x, counties_x, cities_x)
5. Observe which variables have a high correlation coefficient through scatterplot matrix (recommended by Tatum to take out variables that have a value > 0.7)
6. Visualize distributions of y-variables (budget, kWh) and their transformed counterparts (ln_budget,budgetpercapita, ln_budgetpercapita, etc.) to determine who is normal, who is not, and what regressions are appropriate (normal or poisson)
7. Use StepAIC to take into account which variables are picked by this method 
8. Perform poisson regressions
9. Perform normal regressions for budget
10. Perform normal regressions for kWh
11. Create regression results table by compiling all regressions together using stargazer package
12. T-tests on full dataset (only for K12)
1. Spatial code (only for K12 and counties) 

##Setting up Maps

```{r, Start Here for Spatial}

#make a data frame
all_counties<- map_data("county")
# Lot of observations, this stores geometries of county polygons in an inefficient way (there are 3000 counties in US, but way more observations (storing in point, line way))

#visualize
ggplot(data=all_counties)+
  geom_polygon(aes(x = long, y = lat, fill = region, group = group))+ #if you were using a different coord system you would have to change long and lat
  # fill indicates what we want to colored in, in this case we're doing the states (region=state)
  # group is extremely important. There are lots of point values in this data. Group tells R that all of these points belong in a county
  coord_fixed(1.3)+
  guides(fill = FALSE) #if change to fill = any other color, then you would see the outline of all of the counties (county lines are there even when you put FALSE; however we don't see them at this granularity)

```
```{r, spatial}
#subset to California counties
CA_counties<- subset(x = all_counties, subset = region == "california")
# Check subset with: table(CA_counties$region)
```

##K12
```{r, Start}
#read in data for K-12 EE programs 
K12 <- read.csv("k12_counties_3.csv")
```


```{r, K12 transformations}
#log-transform budget, kWh, and enrollment
K12$ln_budget <- log(K12$budget + 1)
K12$ln_kWh <- log(K12$kWh + 1)
K12$ln_enrollment <- log(K12$enrollment + 1)

#round budget and kWh for any poisson attempts
K12$roundedbudget <- round(K12$budget, 0)
K12$roundedkWh <- round(K12$kWh, 0)

#add a binary indicator for any negative binomial attempts 
K12$ybudget <- ifelse(K12$budget>0, 1, 0)
K12$ykWh <- ifelse(K12$kWh>0, 1, 0)

#per student variables
K12$budgetperstudent <- K12$budget/K12$enrollment
K12$lcffperstudent <- K12$lcff/K12$enrollment
K12$kWhperstudent <- K12$kWh/K12$enrollment
K12$incomepercapita <- K12$income/K12$population

K12$ln_budgetperstudent <- log(K12$budgetperstudent + 1)
K12$ln_lcffperstudent <- log(K12$lcffperstudent + 1)
K12$ln_kWhperstudent <- log(K12$kWhperstudent + 1)

K12$roundedlnbudgetper <- round(K12$ln_budgetperstudent, 0)
K12$roundedlnkWhper <- round(K12$ln_kWhperstudent, 0)

K12$roundedlnbudget <- round(K12$ln_budget, 0)
K12$roundedlnkWh <- round(K12$ln_kWh, 0)

K12$roundedbudgetper <- round(K12$budgetperstudent, 0)
K12$roundedkWhper <- round(K12$kWhperstudent, 0)

#scale population and lcff
K12$population_scaled <- K12$population/100000
K12$lcff_scaled <- K12$lcff/100000
K12$lcffperstudent_scaled <- K12$lcffperstudent/100000
K12$income_scaled <- K12$income/100000
```

```{r, K12 subsetting}
#only considering the counties that did participate 
K12_subset <- subset(K12, budget > 0)
#only considering the counties that did not participate 
K12_no = subset(K12, budget = 0)

#quantiles
quantile(K12$ln_enrollment)
#split into groups based ln_enrollment 
K12_1 <- subset(K12, ln_enrollment <= 8.707647)
K12_2 <- subset(K12, ln_enrollment > 8.707647 & ln_enrollment <= 10.336342)
K12_3 <- subset(K12, ln_enrollment > 10.336342 & ln_enrollment <= 11.532680)
K12_4 <- subset(K12, ln_enrollment > 11.532680)

```

```{r, K12 subsetting outliers}
#outliers
#boxplot(K12_subset$budget)

test.K12 <- glm(budget ~ 1, data = K12_subset)
K12CDs <- cooks.distance(test.K12)

# Plot the Cook's Distance using the traditional 4/n criterion
K12_sample_size <- nrow(K12_subset)
plot(K12CDs, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4/K12_sample_size, col="red")  # add cutoff line
text(x=1:length(K12CDs)+1, y=K12CDs, labels=ifelse(K12CDs>4/K12_sample_size, names(K12CDs),""), col="red")  # add labels

#19 (LA) is an outlier
```
```{r, subsetting outliers}
#taking out the outliers 
K12_x <- K12_subset[!(K12_subset$county %in% c("LOS ANGELES")),] 
```


```{r, K12 scatterplot a}
#this gives us correlation coefficients, distributions, and scatterplots
#scatter plot matrix to check for correlation between all variables and investment/energy savings
plot.K12a <- data.frame(K12$kWh, K12$budget, K12$lcff, K12$enrollment, K12$perc_title1, K12$perc_eligible_frpm, K12$perc_eligible_free, K12$income, K12$irr, K12$dac, K12$ces_score, K12$ces_percentile, K12$obf, K12$resource) #subset out these variables
colnames(plot.K12a)<-c("kWh", "budget", "lcff", "enrollment","p_title1", "p_frpm", "p_free", "income", "IRR","dac", "ces_score", "ces_percentile", "obf","resource") #change column names back to original
pairs.panels(plot.K12a, density = TRUE, cor=TRUE, lm=TRUE) #scatterplot
```
```{r, K12 scatterplot b}
#to see the school count variables
plot.K12b <- data.frame(K12$kWh, K12$budget, K12$lcff, K12$enrollment, K12$no_title1, K12$frmp_count, K12$free_meal_count, K12$income, K12$irr, K12$dac, K12$ces_score, K12$ces_percentile, K12$obf, K12$resource) 
colnames(plot.K12b)<-c("kWh", "budget", "lcff", "enrollment","no_title1", "frpm_count", "free_count", "income", "IRR","dac", "ces_score", "ces_percentile", "obf","resource") 
pairs.panels(plot.K12b, density = TRUE, cor=TRUE, lm=TRUE) 
```
```{r, K12 scatterplot c}
#to see the correlation between school count and percentage variables
plot.K12c <- data.frame(K12$kWh, K12$budget,K12$no_title1, K12$frmp_count, K12$free_meal_count, K12$perc_title1, K12$perc_eligible_frpm, K12$perc_eligible_free) 
colnames(plot.K12c)<-c("kWh", "budget", "p_title1", "p_frpm", "p_free","no_title1", "frpm_count", "free_count") 
pairs.panels(plot.K12c, density = TRUE, cor=TRUE, lm=TRUE) 
```
```{r, K12 scatterplot d}
plot.K12d <- data.frame(K12_x$kWh, K12_x$budget, K12_x$lcff, K12_x$enrollment, K12_x$perc_title1, K12_x$perc_eligible_frpm, K12_x$perc_eligible_free, K12_x$income, K12_x$irr, K12_x$dac, K12_x$ces_score, K12_x$ces_percentile, K12_x$obf, K12_x$resource) #subset out these variables
colnames(plot.K12d)<-c("kWh", "budget", "lcff", "enrollment","p_title1", "p_frpm", "p_free", "income", "IRR","dac", "ces_score", "ces_percentile", "obf","resource") #change column names back to original
pairs.panels(plot.K12d, density = TRUE, cor=TRUE, lm=TRUE) #scatterplot
```

```{r, K12 visualize}
#explore our y-variables on full dataset
hist(K12$budget)
hist(K12$kWh)
hist(K12$ln_budget)
hist(K12$ln_kWh)
hist(K12$budgetperstudent)
hist(K12$kWhperstudent)
hist(K12$ln_budgetperstudent)
hist(K12$ln_kWhperstudent)
#explore our y-variables on programs that did participate 
hist(K12_subset$budget)
hist(K12_subset$kWh)
hist(K12_subset$lcff)
hist(K12_subset$ln_budget)
hist(K12_subset$ln_kWh)
hist(K12_subset$budgetperstudent)
hist(K12_subset$kWhperstudent)
hist(K12_subset$ln_budgetperstudent)
hist(K12_subset$ln_kWhperstudent)
```
```{r, K12 visualize}
#explore our y-variables for dataset WITHOUT outliers nor counties that didn't participate
hist(K12_x$budget)
hist(K12_x$kWh)
hist(K12_x$lcff)
hist(K12_x$ln_budget)
hist(K12_x$ln_kWh)
hist(K12_x$budgetperstudent)
hist(K12_x$kWhperstudent)
hist(K12_x$ln_budgetperstudent)
hist(K12_x$ln_kWhperstudent)
```

```{r, K12 Step AIC selection}
#testing these variables: lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income + perc_eligible_frpm + perc_title1

#performs stepwise model selection by AIC for budget
stepAIC(lm(ln_budget ~ lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x))

#ln_budget ~ lcff_scaled + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x
#ln_budgetperstudent ~ lcff_scaled + dac + perc_eligible_frpm + perc_title1, data = K12_x
```

```{r, K12 Step AIC selection}
#performs stepwise model selection by AIC for kWh
stepAIC(lm(ln_kWhperstudent ~ lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x))

#ln_kWh ~ lcff_scaled + lcffperstudent_scaled, data = K12_x
#ln_kWhperstudent ~ lcff_scaled + lcffperstudent_scaled + dac + perc_eligible_frpm + perc_title1, data = K12_x
```

```{r, K12 regression}
#regression from budget

K12.2 <- glm(ln_budget ~  lcff_scaled + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x) #Step AIC picks

K12.4 <- glm(ln_budget ~ lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x) #intuitive picks

K12.6 <- glm( ln_budgetperstudent ~ lcff_scaled + dac + perc_eligible_frpm + perc_title1, data = K12_x) #Step AIC picks; important to note this y variable does not meet normal assumptions

K12.8 <- glm(ln_budgetperstudent ~ lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x) #intuitive picks; important to note this y variable does not meet normal assumptions

stargazer(K12.2, K12.4, K12.6, K12.8, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, K12 regression}
#regression from kWh

K12.1 <- glm(ln_kWh ~  lcff_scaled + lcffperstudent_scaled, data = K12_x) #Step AIC picks

K12.3 <- glm(ln_kWh ~ lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x) #intuitive picks

K12.5 <- glm( ln_kWhperstudent ~ lcff_scaled + lcffperstudent_scaled + dac + perc_eligible_frpm + perc_title1, data = K12_x) #Step AIC picks; important to note this y variable does not meet normal assumptions

K12.7 <- glm(ln_kWhperstudent ~ lcff_scaled + lcffperstudent_scaled + irr + dac + obf + income_scaled + perc_eligible_frpm + perc_title1, data = K12_x) #intuitive picks; important to note this y variable does not meet normal assumptions

stargazer(K12.1, K12.3, K12.5, K12.7, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```


```{r, K12 regression}
#final regression picks BUDGET
#type="text" if you want to see the table in the console 

stargazer(K12.2, K12.4, K12.6, K12.8, type="text", title="Budget Regression Results (K12)", align=TRUE, dep.var.labels=c("Log of Budget", "Log Budget Per Student"), order=c("perc_eligible_frpm", "perc_title1", "lcff_scaled", "lcffperstudent_scaled", "income_scaled", "irr", "dac", "obf"), covariate.labels=c("Percent Eligible for Free or Reduced Meal Plan","Percent of Title 1 Schools", "LCFF (Scaled by 100000)","LCFF per student (Scaled by 100000)","Income (Scaled by 100000)", "IRR","DAC Proportion", "Percent of On-bill Financing"), omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, K12 regression}
#final regression picks KWH

stargazer(K12.1, K12.3, K12.5, K12.7, type="text", title="Energy Savings Regression Results (K12)", align=TRUE, dep.var.labels=c("Log of kWh", "Log kWh Per Student"), order=c("perc_eligible_frpm", "perc_title1", "lcff_scaled", "lcffperstudent_scaled", "income_scaled", "irr", "dac", "obf"), covariate.labels=c("Percent Eligible for Free or Reduced Meal Plan","Percent of Title 1 Schools", "LCFF (Scaled by 100000)","LCFF per student (Scaled by 100000)","Income (Scaled by 100000)", "IRR","DAC Proportion", "Percent of On-bill Financing"), omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, K12 subsetting}
#Create quantiles of budget
quantile(K12$budgetperstudent)

#create top and bottom quantile
K12fbottomq_budget = subset(K12, budgetperstudent <= 0) 
K12ftopq_budget = subset(K12, budgetperstudent > 0.81462460)

#create quantiles of energy savings
quantile(K12$kWhperstudent)

#create top and bottom quantile
K12fbottomq_energy = subset(K12, kWhperstudent <= 0) 
K12ftopq_energy = subset(K12, kWhperstudent > 1.618383175)
```


```{r, K12 t-tests}
#t tests to compare means of irr in top and bottom
t.test(K12fbottomq_budget$irr, K12ftopq_budget$irr, na.action = na.omit) #significant
t.test(K12fbottomq_energy$irr, K12ftopq_energy$irr, na.action = na.omit) #significant

#t tests to compare means of lcff in top and bottom
t.test(K12fbottomq_budget$lcffperstudent, K12ftopq_budget$lcffperstudent) #not sig
t.test(K12fbottomq_energy$lcffperstudent, K12ftopq_energy$lcffperstudent) #not sig

#t tests to compare means of perc_title_1 in top and bottom
t.test(K12fbottomq_budget$perc_title1, K12ftopq_budget$perc_title1) #not sig
t.test(K12fbottomq_energy$perc_title1, K12ftopq_energy$perc_title1) #not sig

#t tests to compare means of perc_frpm in top and bottom
t.test(K12fbottomq_budget$perc_eligible_frpm, K12ftopq_budget$perc_eligible_frpm) #not sig
t.test(K12fbottomq_energy$perc_eligible_frpm, K12ftopq_energy$perc_eligible_frpm) #not sig

#t tests to compare means of no_title_1 in top and bottom
t.test(K12fbottomq_budget$no_title1, K12ftopq_budget$no_title1) 
t.test(K12fbottomq_energy$no_title1, K12ftopq_energy$no_title1) 

#t tests to compare means of dac proportion in top and bottom
t.test(K12fbottomq_budget$dac, K12ftopq_budget$dac) #not sig
t.test(K12fbottomq_energy$dac, K12ftopq_energy$dac) #not sig

#t tests to compare means of ces score median in top and bottom
t.test(K12fbottomq_budget$ces_score, K12ftopq_budget$ces_score)  #not sig
t.test(K12fbottomq_energy$ces_score, K12ftopq_energy$ces_score) #not sig

#compare WA
t.test(K12fbottomq_budget$perc_wa, K12ftopq_budget$perc_wa)  #not sig
t.test(K12fbottomq_energy$perc_wa, K12ftopq_energy$perc_wa)  #not sig

#compare BA
t.test(K12fbottomq_budget$perc_ba, K12ftopq_budget$perc_ba)  #sig
t.test(K12fbottomq_energy$perc_ba, K12ftopq_energy$perc_ba)  #not sig

#compare AA
t.test(K12fbottomq_budget$perc_aa, K12ftopq_budget$perc_aa)  #sig
t.test(K12fbottomq_energy$perc_aa, K12ftopq_energy$perc_aa)  #sig

#compare IA
t.test(K12fbottomq_budget$perc_ia, K12ftopq_budget$perc_ia)  #sig
t.test(K12fbottomq_energy$perc_ia, K12ftopq_energy$perc_ia)  #sig

#compare NA
t.test(K12fbottomq_budget$perc_na, K12ftopq_budget$perc_na)  #not sig
t.test(K12fbottomq_energy$perc_na, K12ftopq_energy$perc_na)  #not sig 

#compare income per capita
t.test(K12fbottomq_budget$income, K12ftopq_budget$income) #sig
t.test(K12fbottomq_energy$income, K12ftopq_energy$income) #sig
```

```{r, K12 Mann-Whitney}
#compare WA
wilcox.test(K12fbottomq_budget$perc_wa, K12ftopq_budget$perc_wa)  #W = 280, p-value = 0.0209 sig
wilcox.test(K12fbottomq_energy$perc_wa, K12ftopq_energy$perc_wa)  #W = 284, p-value = 0.06052 not sig

#compare BA
wilcox.test(K12fbottomq_budget$perc_ba, K12ftopq_budget$perc_ba)  #W = 100, p-value = 0.00935 sig
wilcox.test(K12fbottomq_energy$perc_ba, K12ftopq_energy$perc_ba)  #W = 128, p-value = 0.03676 sig

#compare AA
wilcox.test(K12fbottomq_budget$perc_aa, K12ftopq_budget$perc_aa)  #W = 63, p-value = 0.0001866 sig
wilcox.test(K12fbottomq_energy$perc_aa, K12ftopq_energy$perc_aa)  #W = 83, p-value = 0.0008458 sig

#compare IA
wilcox.test(K12fbottomq_budget$perc_ia, K12ftopq_budget$perc_ia)  #W = 315, p-value = 0.0007997 sig
wilcox.test(K12fbottomq_energy$perc_ia, K12ftopq_energy$perc_ia)  #W = 317, p-value = 0.005653 sig

#compare NA
wilcox.test(K12fbottomq_budget$perc_na, K12ftopq_budget$perc_na)  #W = 92, p-value = 0.004585 sig
wilcox.test(K12fbottomq_energy$perc_na, K12ftopq_energy$perc_na)  #W = 117, p-value = 0.01715 sig
```

```{r, K12 subsetting}
#curious to see how t-tests differ when its just within the group that is participating ((K12 SUBSET))

quantile(K12_subset$budgetperstudent)
K12sbottomq_budget = subset(K12_subset, budgetperstudent <= 0.22256311)
K12stopq_budget = subset(K12_subset, budgetperstudent > 2.87458264)

quantile(K12_subset$kWhperstudent)
K12sbottomq_energy = subset(K12_subset, kWhperstudent <= 0.3356844)
K12stopq_energy = subset(K12_subset, kWhperstudent > 4.4421836)

```

```{r, K12 t-tests}

#compare WA
t.test(K12sbottomq_budget$perc_wa, K12stopq_budget$perc_wa)  
t.test(K12sbottomq_energy$perc_wa, K12stopq_energy$perc_wa) 

#compare BA
t.test(K12sbottomq_budget$perc_ba, K12stopq_budget$perc_ba)  
t.test(K12sbottomq_energy$perc_ba, K12stopq_energy$perc_ba) 

#compare AA
t.test(K12sbottomq_budget$perc_aa, K12stopq_budget$perc_aa)  
t.test(K12sbottomq_energy$perc_aa, K12stopq_energy$perc_aa) 

#compare IA
t.test(K12sbottomq_budget$perc_ia, K12stopq_budget$perc_ia)  
t.test(K12sbottomq_energy$perc_ia, K12stopq_energy$perc_ia) 

#compare NA
t.test(K12sbottomq_budget$perc_na, K12stopq_budget$perc_na)  
t.test(K12sbottomq_energy$perc_na, K12stopq_energy$perc_na) 
```

<!-- ```{r, K12 spatial} -->
<!-- #load data -->
<!-- K12_file   = 'k12_counties_3.csv' -->
<!-- #df <- fread(K12_file) #, colClasses = c('character', df$county_fips)) -->

<!-- K12_df <- fread(K12_file, header = T, colClasses = c('character', rep(NA, 38))) -->
<!-- # best practice is to classify column as a character as you read the data in. -->
<!-- ``` -->

```{r, K12 spatial}
#the county names don't match (our data has capitalized first letters County =/= county), so we have to change it
K12_to_plot <- K12[!(K12$county %in% c("LOS ANGELES")),] # K12 has all counties. K12_x has no outliers or NAs
K12_df_edited <- K12_to_plot # no outlier - no Los Angeles
K12_df_edited$county <- str_to_lower(K12_df_edited$county)
head(K12_df_edited) #check to see if the county names did change
```
```{r, K12 spatial}
#merge on count name
#check dimensions before the merge
dim(K12_df_edited)
dim(CA_counties)

#merge
merged_K12data<- left_join(CA_counties, K12_df_edited, by = c("subregion"="county")) #the left_join takes all the unique values from left one and only add the matches from the right (doesn't matter in this case because they WILL match since we checked already)
dim(merged_K12data)
```

```{r, K12 spatial}

#visualize
a <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `perc_title1`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("Percent of Title 1 Schools")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis() +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#xy plot of budget/kWh vs perc_title1
b <- xyplot(budget ~ perc_title1, data = merged_K12data)
c <- xyplot(kWh ~ perc_title1, data = merged_K12data)

#there are ways to change colors and get rid of the axis (read more into ggplot:http://sape.inf.usi.ch/quick-reference/ggplot2/colour)

```

```{r, K12 spatial}

#visualize
d <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `perc_eligible_frpm`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("Percent of Students Eligible for Free and Reduced Meal Plans")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis() +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#xy plot of budget/kWh vs perc_eligible_frpm
e <- xyplot(budget ~ perc_eligible_frpm, data = merged_K12data)
f <- xyplot(kWh ~ perc_eligible_frpm, data = merged_K12data)
```
```{r, K12 spatial}
#arranging the maps for perc_eligible_frpm and perc_title1
grid.arrange(a, d, ncol=2)

#more information on Grid Arrange and other complex layouts https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html#multiple-pages-output
```


```{r, counties spatial maps data}


# Budget
k1 <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `ln_budget`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Log of Budget \n ")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option = "mako", direction=1, na.value='white', discrete = TRUE) +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Budget per capita
k2 <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `ln_budgetperstudent`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Log of Budget per Capita \n ")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Percent of Students Eligible for Free and Reduced Lunch
k3 <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `perc_eligible_frpm`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Students Eligible for Free \nand Reduced Lunch (%)")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Local Control Funding Formula
k4 <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `lcff_scaled`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Local Control Funding Formula\n('00,000)")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Percent of title 1 schools
k5 <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `perc_title1`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Title 1 Schools (%)\n ")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=-1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                           legend.title = element_blank(),
        legend.position = "right",
                plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Disadvantaged Community (%)
k6 <- ggplot(data = merged_K12data)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `dac`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Disadvantaged Community \nProportion")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=-1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                           legend.title = element_blank(),
                          legend.position = "right",
        plot.title=element_text(size=10),
                           panel.grid = element_blank())

#there are ways to change colors and get rid of the axis (read more into ggplot:http://sape.inf.usi.ch/quick-reference/ggplot2/colour)


#arranging the maps for irr
grid.arrange(k1, k3, k4, k2, k5, k6, ncol=3)
k = arrangeGrob(k1, k3, k4, k2, k5, k6, ncol=3)
ggsave(file = "maps_K12_county_analysis.png", k)
```



##Counties

```{r, Start Counties}
#read in data for K-12 EE programs 
counties <- read.csv("counties_3.csv")
```


```{r, counties transformations}
#log-transform 
counties$ln_kWh <- log(counties$kWh + 1)
counties$ln_rev <- log(counties$revenue)
counties$ln_population <- log(counties$population)
counties$ln_budget <- log(counties$budget + 1)

#round budget and kWh
counties$roundedbudget <- round(counties$budget, 0)
counties$roundedkWh <- round(counties$kWh, 0)

#add a binary indicator
counties$ybudget <- ifelse(counties$budget>0, 1, 0)
counties$ykWh <- ifelse(counties$kWh>0, 1, 0)

#per person
counties$budgetpercapita <- counties$budget/counties$population
counties$revpercapita <- counties$revenue/counties$population
counties$kWhpercapita <- counties$kWh/counties$population

counties$ln_budgetpercapita <- log(counties$budgetpercapita + 1)
counties$ln_revpercapita <- log(counties$revpercapita + 1)
counties$ln_kWhpercapita <- log(counties$kWhpercapita + 1)

#round 
counties$roundedbudgetper <- round(counties$budgetpercapita, 0)
counties$roundedkWhper <- round(counties$kWhpercapita, 0)

#scaling
counties$revenue_scaled <- counties$revenue/100000
counties$population_scaled <- counties$population/100000
counties$income_scaled <- counties$income/100000
counties$revpercapita_scaled <- counties$revpercapita/100000
```

```{r, counties subsetting}

#only considering the counties that did participate 
counties_subset <- subset(counties, budget > 0)
#only considering the counties that did not participate 
counties_no = subset(counties, budget = 0)

#quantiles
quantile(counties$ln_population)
#split into groups based ln_enrollment 
counties_1 <- subset(counties, population <= 10.768725)
counties_2 <- subset(counties, population > 10.768725 & population <= 12.125893)
counties_3 <- subset(counties, population > 12.125893 & population <= 13.43974)
counties_4 <- subset(counties, population > 13.43974)
```

```{r, counties subsetting outliers}
#outliers
#boxplot(counties_subset$budget)

test.counties <- glm(budget ~ 1, data = counties_subset)
countiesCDs <- cooks.distance(test.counties)

# Plot the Cook's Distance using the traditional 4/n criterion
counties_sample_size <- nrow(counties_subset)
plot(countiesCDs, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4/counties_sample_size, col="red")  # add cutoff line
text(x=1:length(countiesCDs)+1, y=countiesCDs, labels=ifelse(countiesCDs>4/counties_sample_size, names(countiesCDs),""), col="red")  # add labels

#19 (LA), #37 (SD), and #38 (SF) are outliers
```
```{r, counties subsetting outliers}
#taking out the outliers 
counties_x <- counties_subset[!(counties_subset$county %in% c("LOS ANGELES", "SAN DIEGO", "SAN FRANCISCO")),]
```


```{r, counties scatterplot}
plot.counties_x <- data.frame(counties_x$kWh, counties_x$budget, counties_x$income, counties_x$irr, counties_x$dac, counties_x$ces_score, counties_x$ces_percentile, counties_x$obf, counties_x$revenue, counties_x$revpercapita, counties_x$population) 
colnames(plot.counties_x)<-c("kWh", "budget", "income", "irr", "dac", "ces_score", "ces_percentile", "obf", "revenue", "revenuepercap", "population") 
pairs.panels(plot.counties_x, density = TRUE, cor=TRUE, lm=TRUE) 
```

```{r, counties visualize}
#explore our y-variables 
hist(counties$budget)
hist(counties$kWh)
hist(counties$ln_budget)
hist(counties$ln_kWh)
hist(counties$budgetpercapita)
hist(counties$kWhpercapita)
hist(counties$ln_budgetpercapita)
hist(counties$ln_kWhpercapita)
```
```{r, counties visualize}
hist(counties_subset$budget)
hist(counties_subset$kWh)
hist(counties_subset$revenue)
hist(counties_subset$ln_budget)
hist(counties_subset$ln_kWh)
hist(counties_subset$budgetpercapita)
hist(counties_subset$kWhpercapita)
hist(counties_subset$ln_budgetpercapita)
hist(counties_subset$ln_kWhpercapita)
```

```{r, counties visualize}
hist(counties_x$budget)
hist(counties_x$kWh)
hist(counties_x$revenue)
hist(counties_x$ln_budget)
hist(counties_x$ln_kWh)
hist(counties_x$budgetpercapita)
hist(counties_x$kWhpercapita)
hist(counties_x$ln_budgetpercapita)
hist(counties_x$ln_kWhpercapita)
```

```{r, County Step AIC selection}

#performs stepwise model selection by AIC for budget
stepAIC(lm(ln_budget ~ revenue_scaled + revpercapita_scaled + dac + irr + population_scaled + obf + income_scaled, data = counties_x))

#ln_budgetpercapita ~ revpercapita_scaled + population_scaled + obf, data = counties_x
#ln_budget ~ revenue_scaled + revpercapita_scaled + irr + population_scaled + obf, data = counties_x

```

```{r, County Step AIC selection}

#performs stepwise model selection by AIC for kWh
stepAIC(lm(ln_kWh ~ revenue_scaled + revpercapita_scaled + dac + irr + population_scaled + obf + income_scaled, data = counties_x))

#ln_kWhpercapita ~ revpercapita_scaled + irr + population_scaled, data = counties_x
#ln_kWh ~ revenue_scaled + revpercapita_scaled + irr + population_scaled, data = counties_x
```

```{r, counties regression}
#regression from budget

counties.2 <- glm(ln_budget ~ revenue_scaled + revpercapita_scaled + irr + population_scaled + obf, data = counties_x) #Step AIC picks

counties.4 <- glm(ln_budget ~ revenue_scaled + revpercapita_scaled + dac + irr + population_scaled + obf + income_scaled, data = counties_x) #intuitive picks

counties.6 <- glm(ln_budgetpercapita ~ revpercapita_scaled + population_scaled + obf, data = counties_x) #Step AIC picks

counties.8 <- glm(ln_budgetpercapita ~ revenue_scaled + revpercapita_scaled + dac + irr + population_scaled + obf + income_scaled, data = counties_x) #intuitive picks

stargazer(counties.2, counties.4, counties.6, counties.8, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, counties regression}
#regression from kWh

counties.1 <- glm(ln_kWh ~  revenue_scaled + revpercapita_scaled + irr + population_scaled, data = counties_x) #Step AIC picks

counties.3 <- glm(ln_kWh ~ revenue_scaled + revpercapita_scaled + dac + irr + population_scaled + obf + income_scaled, data = counties_x) #intuitive picks

counties.5 <- glm(ln_kWhpercapita ~ revpercapita_scaled + irr + population_scaled, data = counties_x) #Step AIC picks

counties.7 <- glm(ln_kWhpercapita ~ revenue_scaled + revpercapita_scaled + dac + irr + population_scaled + obf + income_scaled, data = counties_x) #intuitive picks

stargazer(counties.1, counties.3, counties.5, counties.7, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, counties regression}
#final regression picks BUDGET
#to export table to nice format -> type="html", out= "tablecounty1.htm", title="Budget Regression......

stargazer(counties.2, counties.4, counties.6, counties.8, type="text", title="Budget Regression Results (County)", align=TRUE, dep.var.labels=c("Log of Budget", "Log Budget Per Capita"), order=c("revenue_scaled", "revpercapita_scaled", "income_scaled", "irr", "dac", "obf", "population_scaled"), covariate.labels=c("Revenue (Scaled by 100000)", "Revenue per capita (Scaled by 100000)", "Income (Scaled by 100000)","IRR","DAC Proportion", "Percent of On-bill Financing","Population (Scaled by 100000)"), omit.stat=c("LL","ser","f"), no.space=TRUE)
```
```{r, counties regression}
#final regression picks KWH

stargazer(counties.1, counties.3, counties.5, counties.7, type="text", title="Energy Savings Regression Results (County)", align=TRUE, dep.var.labels=c("Log of kWh", "Log kWh Per Capita"), order=c("revenue_scaled", "revpercapita_scaled", "income_scaled", "irr", "dac", "obf", "population_scaled"), covariate.labels=c("Revenue (Scaled by 100000)", "Revenue per capita (Scaled by 100000)", "Income (Scaled by 100000)","IRR","DAC Proportion", "Percent of On-bill Financing","Population (Scaled by 100000)"), omit.stat=c("LL","ser","f"), no.space=TRUE)
```
```{r, counties subsetting}

#Create quantiles of budget
quantile(counties$budgetpercapita)

#create top and bottom quantile
countiesfbottomq_budget = subset(counties, budgetpercapita <= 1.528228)
countiesftopq_budget = subset(counties, budgetpercapita > 7.940744)

#create quantiles of energy savings
quantile(counties$kWhpercapita)

#create top and bottom quantile
countiesfbottomq_energy = subset(counties, kWhpercapita <= 1.377546)
countiesftopq_energy = subset(counties, kWhpercapita > 11.231349)
```

```{r}
#t tests to compare means of irr in top and bottom
t.test(bottomq_budget$IRR, topq_budget$IRR, na.action = na.omit) #super sig--> makes sense because only real significance in regression
t.test(bottomq_energy$IRR, topq_energy$IRR, na.action = na.omit, var.equal = TRUE) #significant

#t tests to compare means of budget per capita in top and bottom
t.test(bottomq_energy$budget_per_cap, topq_energy$budget_per_cap) #significant -> this more important cuz shoes relationship between energy and budget

#t tests to compare means of total revenue in top and bottom
t.test(bottomq_budget$TotalRevenue, topq_budget$TotalRevenue) #significant
t.test(bottomq_energy$TotalRevenue, topq_energy$TotalRevenue) #not significant

#t tests to compare means of population in top and bottom
t.test(bottomq_budget$Population, topq_budget$Population) #significant
t.test(bottomq_energy$Population, topq_energy$Population) #not significant, not different 

#t tests to compare means of dac proportion in top and bottom
t.test(bottomq_budget$dac_proportion, topq_budget$dac_proportion, var.equal = TRUE) #not significant--> not significant in regressions
t.test(bottomq_energy$dac_proportion, topq_energy$dac_proportion, var.equal = TRUE) #not significant

#t tests to compare means of ces score median in top and bottom
t.test(bottomq_budget$ces_score_median, topq_budget$ces_score_median, var.equal = TRUE) #not significant --> makes sense because not significant with budget in regression
t.test(bottomq_energy$ces_score_median, topq_energy$ces_score_median) #not significant

#t tests to compare means of mean income in top and bottom
t.test(bottomq_budget$Mean.Income, topq_budget$Mean.Income, var.equal = TRUE) #not sig --> makes sense because not significant in any of the regressions
t.test(bottomq_energy$Mean.Income, topq_energy$Mean.Income, var.equal = TRUE) #not sig --> makes sense because not significant in any of the regressions
```

```{r}
#look at average percentage of certain demographics within top and bottom quantile based on budget per cap
t.test(bottomq_budget$perc_wa_avg,topq_budget$perc_wa_avg)

t.test(bottomq_budget$perc_ba_avg, topq_budget$perc_ba_avg)

t.test(bottomq_budget$perc_aa_avg, topq_budget$perc_aa_avg)

t.test(bottomq_budget$perc_ia_avg, topq_budget$perc_ia_avg)

t.test(bottomq_budget$perc_na_avg, topq_budget$perc_na_avg, var.equal = TRUE)
```

```{r}
#look at average percentage of certain demographics within top and bottom quantile based on kwh per cap
t.test(bottomq_energy$perc_wa_avg, topq_energy$perc_wa_avg)

t.test(bottomq_energy$perc_ba_avg, topq_energy$perc_ba_avg)

t.test(bottomq_energy$perc_aa_avg, topq_energy$perc_aa_avg)

t.test(bottomq_energy$perc_ia_avg, topq_energy$perc_ia_avg)

t.test(bottomq_energy$perc_na_avg, topq_energy$perc_na_avg)
```

```{r, counties Mann-Whitney}
#compare WA
wilcox.test(countiesfbottomq_budget$perc_wa, countiesftopq_budget$perc_wa)  #W = 99, p-value = 0.5949 not sig
wilcox.test(countiesfbottomq_energy$perc_wa, countiesftopq_energy$perc_wa)  #W = 107, p-value = 0.8381 not sig

#compare BA
wilcox.test(countiesfbottomq_budget$perc_ba, countiesftopq_budget$perc_ba)  #W = 103, p-value = 0.713 not sig
wilcox.test(countiesfbottomq_energy$perc_ba, countiesftopq_energy$perc_ba)  #W = 89, p-value = 0.3453 not sig

#compare AA
wilcox.test(countiesfbottomq_budget$perc_aa, countiesftopq_budget$perc_aa)  #W = 86, p-value = 0.2854 not sig
wilcox.test(countiesfbottomq_energy$perc_aa, countiesftopq_energy$perc_aa)  #W = 71, p-value = 0.08919 not sig

#compare IA
wilcox.test(countiesfbottomq_budget$perc_ia, countiesftopq_budget$perc_ia)  #W = 138, p-value = 0.3046 not sig
wilcox.test(countiesfbottomq_energy$perc_ia, countiesftopq_energy$perc_ia)  #W = 152, p-value = 0.1064 not sig

#compare NA
wilcox.test(countiesfbottomq_budget$perc_na, countiesftopq_budget$perc_na)  #W = 103, p-value = 0.713 not sig
wilcox.test(countiesfbottomq_energy$perc_na, countiesftopq_energy$perc_na)  #W = 95, p-value = 0.4864 not sig
```

# ```{r, counties spatial}
# #load data
# counties_file   = 'counties_3.csv'
# #df <- fread(K12_file) #, colClasses = c('character', df$county_fips))
# 
# counties_df <- fread(counties_file, header = T, colClasses = c('character', rep(NA, 31)))
# ```

```{r, counties spatial}
#the county names don't match (our data has capitalized first letters County =/= county), so we have to change it
counties_to_include = counties_x[!(counties_x$county %in% c("LOS ANGELES", "SAN DIEGO", "SAN FRANCISCO")),] # Use counties for all; counties_x to exclude zero budgets and outliers. 
counties_df_edited <- counties_to_include # no outliers - LA, SF, SD

counties_df_edited$county <- str_to_lower(counties_df_edited$county)
head(counties_df_edited) #check to see if the county names did change
```
```{r, counties spatial}
#merge on count name
#check dimensions before the merge
dim(counties_df_edited)
dim(CA_counties)
```
```{r, counties spatial}
#merge
merged_countiesdata<- left_join(CA_counties, counties_df_edited, by = c("subregion"="county")) 
dim(merged_countiesdata)
```
```{r, counties spatial}

#visualize
g <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `irr`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("Rurality by County")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis() +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#xy plot of budget/kWh vs perc_title1
h <- xyplot(budget ~ irr, data = merged_countiesdata)
i <- xyplot(kWh ~ irr, data = merged_countiesdata)

#there are ways to change colors and get rid of the axis (read more into ggplot:http://sape.inf.usi.ch/quick-reference/ggplot2/colour)

```

```{r, counties spatial}

#LOG Tax Revenue
merged_countiesdata$log_revenue <- log(merged_countiesdata$revenue + 1)

#visualize
j <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `log_revenue`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("Revenue (Log) by County")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis() +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#xy plot of budget/kWh vs perc_eligible_frpm
k <- xyplot(budget ~ revenue, data = merged_countiesdata)
l <- xyplot(kWh ~ revenue, data = merged_countiesdata)
```

```{r, counties spatial}
#arranging the maps for irr
grid.arrange(g, h, i, ncol=2)
```
```{r, counties spatial}

#Budget
merged_countiesdata$ln_budget <- log(merged_countiesdata$budget + 1)

#visualize
ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `ln_budget`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("budget (Log) by County")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis() +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#xy plot
xyplot(budget ~ dac, data = merged_countiesdata)
xyplot(kWh ~ budget, data = merged_countiesdata)
```

```{r, counties spatial}

#kWh
merged_countiesdata$ln_kWh <- log(merged_countiesdata$kWh + 1)

#visualize
ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `ln_kWh`))+ #change the fill argument from plot to plot
  coord_fixed(1.3)+
  ggtitle("kWh (Log) by County")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis() +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           panel.grid = element_blank())

#xy plot 
xyplot(kWh ~ dac, data = merged_countiesdata)
xyplot(kWh ~ budget, data = merged_countiesdata)
```

```{r, counties spatial maps data}

# Budget
g1 <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `ln_budget`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Log of Budget\n ")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option = "mako", direction=1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
                           legend.position = "right",
                          plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Budget per capita
g2 <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `ln_budgetpercapita`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Log of Budget per capita\n ")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                  plot.title=element_text(size=10),
                           panel.grid = element_blank())

# IRR
g3 <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `irr`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Rurality (Index of \nRelative Rurality)")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=-1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                  plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Revenue per Capita
g4 <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `revpercapita`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Revenue per Capita\n(USD)")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=-1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                          legend.title = element_blank(),
        legend.position = "right",
                  plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Population
g5 <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `population_scaled`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Population \n('00,000)")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=-1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                           legend.title = element_blank(),
        legend.position = "right",
                  plot.title=element_text(size=10),
                           panel.grid = element_blank())

# Percent of on bill financing
g6 <- ggplot(data = merged_countiesdata)+
  geom_polygon(aes(x = long, y = lat, group = group, fill= `obf`), color='grey', size=0.1)+ 
  coord_fixed(1.3)+
  ggtitle("Percent of On-bill \nFinancing")+ #change title from plot to plot
  theme_classic()+ #order matters in ggplot
  scale_fill_viridis(option="mako", direction=1, na.value='white') +
  theme(axis.title.x = element_blank(),
                           axis.title.y = element_blank(),
                           axis.text = element_blank(),
                           axis.ticks = element_blank(),
                           axis.line = element_blank(),
                           legend.title = element_blank(),
                          legend.position = "right",
                  plot.title=element_text(size=10),
                           panel.grid = element_blank())

#there are ways to change colors and get rid of the axis (read more into ggplot:http://sape.inf.usi.ch/quick-reference/ggplot2/colour)


#arranging the maps 
grid.arrange(g1, g3, g4, g2, g5, g6, ncol=3)
g = arrangeGrob(g1, g3, g4, g2, g5, g6, ncol=3)
ggsave(file = "maps_local_govt_county_analysis_no_NAs.png", g)
```

```{r, local govt and K12 maps}
#arranging local government and K12 maps
grid.arrange(g1, g3, g4, g2, g5, g6, k1, k2, k3, k4, k5, k6, ncol=2)
```

##Cities

```{r, Start Cities}
cities <- read.csv("cities_3.csv")
```

```{r, cities transformations}
#log-transform 
cities$ln_budget <- log(cities$budget + 1)
cities$ln_kWh <- log(cities$kWh + 1)
cities$ln_rev <- log(cities$revenue)

#round budget and kWh
cities$roundedbudget <- round(cities$budget, 0)
cities$roundedkWh <- round(cities$kWh, 0)

#add a binary indicator
cities$ybudget <- ifelse(cities$budget>0, 1, 0)
cities$ykWh <- ifelse(cities$kWh>0, 1, 0)

#per student
cities$budgetpercapita <- cities$budget/cities$population
cities$revpercapita <- cities$revenue/cities$population
cities$kWhpercapita <- cities$kWh/cities$population

cities$ln_budgetpercapita <- log(cities$budgetpercapita)
cities$ln_revpercapita <- log(cities$revpercapita)
cities$ln_kWhpercapita <- log(cities$kWhpercapita + 1)

#round budget and kWh
cities$roundedbudgetper <- round(cities$budgetpercapita, 0)
cities$roundedkWhper <- round(cities$kWhpercapita, 0)

#scaling
cities$population_scaled <- cities$population/100000
cities$revenue_scaled <- cities$revenue/100000
cities$revpercapita_scaled <- cities$revpercapita/100000
cities$income_scaled <- cities$income/100000
```

```{r, city subsetting}
#get rid of cities with NAs for tax revenue
cities_na <- na.omit(cities)
```

```{r, cities subsetting outliers}
#outliers
#boxplot(cities_na$budget)

test.cities <- glm(budget ~ 1, data = cities_na)
citiesCDs <- cooks.distance(test.cities)

# Plot the Cook's Distance using the traditional 4/n criterion
cities_sample_size <- nrow(cities_na)
plot(citiesCDs, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4/cities_sample_size, col="red")  # add cutoff line
text(x=1:length(citiesCDs)+1, y=citiesCDs, labels=ifelse(citiesCDs>4/cities_sample_size, names(citiesCDs),""), col="red")  # add labels

#326 (SAN FRANCISCO) and 325 (SAN DIEGO) are outliers and 201 (LA) and 26 (BAKERSFIELD)
```

```{r, cities subsetting}
#taking out the outliers 
cities_x <- cities_na[!(cities_na$city %in% c("SAN DIEGO", "SAN FRANCISCO", "BAKERSFIELD", "LOS ANGELES")),] 
```

```{r, cities scatterplot}
plot.cities_x <- data.frame(cities_x$kWh, cities_x$budget, cities_x$income_scaled, cities_x$ruca, cities_x$dac, cities_x$obf, cities_x$revenue_scaled, cities_x$population, cities_x$revpercapita_scaled) 
colnames(plot.cities_x)<-c("kWh", "budget", "income", "ruca", "dac", "obf", "revenue", "population", "revpercap") 
pairs.panels(plot.cities_x, density = TRUE, cor=TRUE, lm=TRUE) 
```

```{r, cities visualize}
#explore our y-variables for full cities dataset
hist(cities$budget)
hist(cities$kWh)
hist(cities$ln_budget)
hist(cities$ln_kWh)
hist(cities$budgetpercapita)
hist(cities$kWhpercapita)
hist(cities$ln_budgetpercapita)
hist(cities$ln_kWhpercapita)
#explore our y-variables for just cities that had tax revenue
hist(cities_na$budget)
hist(cities_na$kWh)
hist(cities_na$ln_budget)
hist(cities_na$ln_kWh)
hist(cities_na$budgetpercapita)
hist(cities_na$kWhpercapita)
hist(cities_na$ln_budgetpercapita)
hist(cities_na$ln_kWhpercapita)
```
```{r, cities visualize}
#explore our y-variables for dataset WITHOUT outliers nor cities that didn't have tax revenue information 
hist(cities_x$budget)
hist(cities_x$kWh)
hist(cities_x$ln_budget)
hist(cities_x$ln_kWh)
hist(cities_x$budgetpercapita)
hist(cities_x$kWhpercapita)
hist(cities_x$ln_budgetpercapita)
hist(cities_x$ln_kWhpercapita)
```

```{r, City Step AIC selection}

#performs stepwise model selection by AIC for budget
#all variables: revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled 
stepAIC(lm(ln_budgetpercapita ~ revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled, data = cities_x))

#ln_budget ~ revenue_scaled + obf, data = cities_x)
#ln_budgetpercapita ~ revenue_scaled + revpercapita_scaled + ruca + population_scaled + obf, data = cities_x)
```

```{r, City Step AIC selection}

#performs stepwise model selection by AIC for kWh
stepAIC(lm(ln_kWhpercapita ~ revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled, data = cities_x))

#ln_kWh ~ revenue_scaled + obf, data = cities_x)
#ln_kWhpercapita ~ revenue_scaled + ruca + dac + population_scaled + obf, data = cities_x)
```

```{r, cities regression}
#regression from budget

cities.2 <- glm(ln_budget ~  revenue_scaled + obf, data = cities_x) #Step AIC picks

cities.4 <- glm(ln_budget ~ revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled, data = cities_x) #intuitive picks

cities.6 <- glm(ln_budgetpercapita ~ revenue_scaled + revpercapita_scaled + ruca + population_scaled + obf, data = cities_x) #Step AIC picks

cities.8 <- glm(ln_budgetpercapita ~ revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled, data = cities_x) #intuitive picks

stargazer(cities.2, cities.4, cities.6, cities.8, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, cities regression}
#regression from kWh

cities.1 <- glm(ln_kWh ~ revenue_scaled + obf, data = cities_x) #Step AIC picks

cities.3 <- glm(ln_kWh ~ revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled, data = cities_x) #intuitive picks

cities.5 <- glm(ln_kWhpercapita ~ revenue_scaled + ruca + dac + population_scaled + obf, data = cities_x) #Step AIC picks

cities.7 <- glm(ln_kWhpercapita ~ revenue_scaled + revpercapita_scaled + ruca + dac + population_scaled + obf + income_scaled, data = cities_x) #intuitive picks

stargazer(cities.1, cities.3, cities.5, cities.7, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, cities regression}
#final regression picks BUDGET

stargazer(cities.2, cities.4, cities.6, cities.8, type="html", out= "tablecity5.htm", title="Budget Regression Results (Cities)", align=TRUE, dep.var.labels=c("Log of Budget", "Log Budget Per Capita"), order=c("revenue_scaled","revpercapita_scaled", "income_scaled", "ruca", "dac", "obf", "population_scaled"), covariate.labels=c("Revenue (Scaled by 100000)", "Revenue per Capita (Scaled by 100000)","Income (Scaled by 100000)", "RUCA","DAC Proportion", "Percent of On-bill Financing", "Population (Scaled by 100000)"), omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, cities regression}
#final regression picks KWH

stargazer(cities.1, cities.3, cities.5, cities.7, type="html", out= "tablecity4.htm", title="Energy Savings Regression Results (Cities)", align=TRUE, dep.var.labels=c("Log of kWh", "Log kWh Per Capita"), order=c("revenue_scaled","revpercapita_scaled", "income_scaled", "ruca", "dac", "obf", "population_scaled"), covariate.labels=c("Revenue (Scaled by 100000)", "Revenue per Capita (Scaled by 100000)","Income (Scaled by 100000)", "RUCA","DAC Proportion", "Percent of On-bill Financing", "Population (Scaled by 100000)"), omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, cities subsetting}
#Create quantiles of budget
quantile(cities$budgetpercapita)

#create top and bottom quantile
citiesfbottomq_budget = subset(cities, budgetpercapita <= 1.292106)
citiesftopq_budget = subset(cities, budgetpercapita > 8.699818)

#create quantiles of energy savings
quantile(cities$kWhpercapita)

#create top and bottom quantile
citiesfbottomq_energy = subset(cities, kWhpercapita <= 1.141914)
citiesftopq_energy = subset(cities, kWhpercapita > 13.167434)
```

```{r, cities t-tests}

#t tests to compare means of dac proportion in top and bottom
t.test(citiesfbottomq_budget$dac, citiesftopq_budget$dac, var.equal = TRUE) 
t.test(citiesfbottomq_energy$dac, citiesftopq_energy$dac, var.equal = TRUE) 

#t tests to compare means of ces score median in top and bottom
t.test(citiesfbottomq_budget$ces_score, citiesftopq_budget$ces_score, var.equal = TRUE)  
t.test(citiesfbottomq_energy$ces_score, citiesftopq_energy$ces_score, var.equal = TRUE) 

#t tests to compare means of ruca in top and bottom
t.test(citiesfbottomq_budget$ruca, citiesftopq_budget$ruca, var.equal = TRUE) 
t.test(citiesfbottomq_energy$ruca, citiesftopq_energy$ruca) 

#t tests to compare means of population in top and bottom
t.test(citiesfbottomq_budget$population, citiesftopq_budget$population) #significant
t.test(citiesfbottomq_energy$population, citiesftopq_energy$population) #significant

#t tests to compare means of revenue in top and bottom
t.test(citiesfbottomq_budget$revenue, citiesftopq_budget$revenue) 
t.test(citiesfbottomq_budget$revenue, citiesftopq_budget$revenue) 

#t tests to compare means of ces score median in top and bottom
t.test(citiesfbottomq_budget$income, citiesftopq_budget$income, var.equal = TRUE)  
t.test(citiesfbottomq_energy$income, citiesftopq_energy$income, var.equal = TRUE) 
```

```{r}
#t tests to compare means of dac proportion in top and bottom
t.test(citiesfbottomq_budget$dac, citiesftopq_budget$dac) 
t.test(citiesfbottomq_energy$dac, citiesftopq_energy$dac) 

#t tests to compare means of ces score median in top and bottom
t.test(citiesfbottomq_budget$ces_score, citiesftopq_budget$ces_score)  
t.test(citiesfbottomq_energy$ces_score, citiesftopq_energy$ces_score) 
```


```{r, cities t-tests}
#compare WA
t.test(citiesfbottomq_budget$perc_wa, citiesftopq_budget$perc_wa, var.equal = TRUE)  
t.test(citiesfbottomq_energy$perc_wa, citiesftopq_energy$perc_wa, var.equal = TRUE) 

#compare BA
t.test(citiesfbottomq_budget$perc_ba, citiesftopq_budget$perc_ba, var.equal = TRUE)  
t.test(citiesfbottomq_energy$perc_ba, citiesftopq_energy$perc_ba) 

#compare AA
t.test(citiesfbottomq_budget$perc_aa, citiesftopq_budget$perc_aa, var.equal= TRUE)  
t.test(citiesfbottomq_energy$perc_aa, citiesftopq_energy$perc_aa) 

#compare IA
t.test(citiesfbottomq_budget$perc_ia, citiesftopq_budget$perc_ia, var.equal= TRUE)  
t.test(citiesfbottomq_energy$perc_ia, citiesftopq_energy$perc_ia, var.equal = TRUE) 

#compare NA
t.test(citiesfbottomq_budget$perc_na, citiesftopq_budget$perc_na, var.equal = TRUE)  
t.test(citiesfbottomq_energy$perc_na, citiesftopq_energy$perc_na) 
```

```{r, cities Mann-Whitney}
#compare WA
wilcox.test(citiesfbottomq_budget$perc_wa, citiesftopq_budget$perc_wa)  #W = 5356, p-value = 0.7231 not sig
wilcox.test(citiesfbottomq_energy$perc_wa, citiesftopq_energy$perc_wa)  #W = 5043, p-value = 0.3407 not sig

#compare BA
wilcox.test(citiesfbottomq_budget$perc_ba, citiesftopq_budget$perc_ba)  #W = 5838, p-value = 0.4604 not sig
wilcox.test(citiesfbottomq_energy$perc_ba, citiesftopq_energy$perc_ba)  #W = 6265, p-value = 0.06571 not sig

#compare AA
wilcox.test(citiesfbottomq_budget$perc_aa, citiesftopq_budget$perc_aa)  #W = 6006, p-value = 0.2628 not sig
wilcox.test(citiesfbottomq_energy$perc_aa, citiesftopq_energy$perc_aa)  #W = 6324.5, p-value = 0.0481 sig

#compare IA
wilcox.test(citiesfbottomq_budget$perc_ia, citiesftopq_budget$perc_ia)  #W = 5534.5, p-value = 0.961 not sig
wilcox.test(citiesfbottomq_energy$perc_ia, citiesftopq_energy$perc_ia)  #W = 5166, p-value = 0.5019 not sig

#compare NA
wilcox.test(citiesfbottomq_budget$perc_na, citiesftopq_budget$perc_na)  #W = 6101.5, p-value = 0.1762 not sig
wilcox.test(citiesfbottomq_energy$perc_na, citiesftopq_energy$perc_na)  #W = 5496, p-value = 0.9347 not sig
```

##MISC. K12 Attempts 

```{r, K12 regression}
par(mfrow=c(2,2))
plot(K12.3)
res.K12.3 <- K12.3$residuals#fetching residuals 
qqPlot(K12.3)
shapiro.test(res.K12.3) #seeing if data is normal
summary(K12.3)
```
```{r, K12 model validation}
exp(summary(K12.8)$coefficients[,1])
K12$predict_b <- predict(K12.8, K12, type = "response")

chisq.test(K12$budget, K12$predict_b) 
```

```{r, K12 model validation}
plot(K12.8)
plot(fitdist(K12_x$roundedbudgetper,"pois"))
```

```{r, K12 regression}
#linear regression 

K12x_1.glm <- glm(ln_budgetperstudent ~ lcff + no_title1 + irr + dac + ces_percentile + ces_score + frmp_count + free_meal_count + obf + income + perc_eligible_free + perc_eligible_frpm + number_of_schools + perc_title1 + enrollment + resource, data = K12_x) #all variables

K12x_2.glm <- glm(ln_budgetperstudent ~ lcffperstudent + irr + dac + obf + income + perc_eligible_frpm + perc_title1 + resource, data = K12_x) #variables I am curious about + perc

K12x_3.glm <- glm(ln_budgetperstudent ~ ces_percentile + ces_score + perc_eligible_free + perc_eligible_frpm + perc_title1, data = K12_x) #StepAIC picks 

K12x_4.glm <- glm(ln_budgetperstudent ~ lcff + no_title1 + irr + dac + frmp_count + obf + income + resource, data = K12_x) #variables I am curious about + counts
```

```{r, K12 regression}
summary(K12x_1.glm) 
summary(K12x_2.glm) 
summary(K12x_3.glm) #has the smallest AIC and has the most conclusive results!!!!
summary(K12x_4.glm) 

#anova(counties_2f.glm, counties_2g.glm, counties_2h.glm, counties_2i.glm, counties_2j.glm, test="Chisq") #lower p-value means it explains the variance more which is best
AIC(K12x_1.glm, K12x_2.glm, K12x_3.glm, K12x_4.glm)
BIC(K12x_1.glm, K12x_2.glm, K12x_3.glm, K12x_4.glm)
```

```{r, K12 regression results}
stargazer(K12x_2.glm, K12x_3.glm, K12x_4.glm, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, K12 regression}
#linear regression 

K12x_a.glm <- glm(ln_budget ~ lcff + no_title1 + irr + dac + ces_percentile + ces_score + frmp_count + free_meal_count + obf + income + perc_eligible_free + perc_eligible_frpm + number_of_schools + perc_title1 + enrollment + resource, data = K12_x) #all variables

K12x_b.glm <- glm(ln_budget ~ lcff + irr + dac + obf + income + perc_eligible_frpm + number_of_schools + perc_title1, data = K12_x) #variables I am curious about + perc

K12x_c.glm <- glm(ln_budget ~ no_title1 + frmp_count + free_meal_count +  perc_eligible_frpm + number_of_schools + perc_title1, data = K12_x) #StepAIC picks 

K12x_d.glm <- glm(ln_budget ~ lcff + no_title1 + irr + dac + frmp_count + obf + income + number_of_schools + resource, data = K12_x) #variables I am curious about + counts
```

```{r, K12 regression}
summary(K12x_a.glm) 
summary(K12x_b.glm) 
summary(K12x_c.glm) #has the smallest AIC and has the most conclusive results!!!!
summary(K12x_d.glm) 

AIC(K12x_a.glm, K12x_b.glm, K12x_c.glm, K12x_d.glm)
BIC(K12x_a.glm, K12x_b.glm, K12x_c.glm, K12x_d.glm)
```

```{r, K12 regression}
#linear regression

K12x_5.glm <- glm(ln_kWhperstudent ~ lcff + no_title1 + irr + dac + ces_percentile + ces_score + frmp_count + free_meal_count + obf + income + perc_eligible_free + perc_eligible_frpm + number_of_schools + perc_title1 + enrollment + resource, data = K12_x) #all variables

K12x_6.glm <- glm(ln_kWhperstudent ~ lcff + irr + dac + obf + income + perc_eligible_frpm + number_of_schools + perc_title1 + resource, data = K12_x) #variables I am curious about + perc

K12x_7.glm <- glm(ln_kWhperstudent ~ lcff + ces_percentile + ces_score + frmp_count + free_meal_count + number_of_schools + perc_title1, data = K12_x) #StepAIC picks 

K12x_8.glm <- glm(ln_kWhperstudent ~ lcff + no_title1 + irr + dac + frmp_count + obf + income + number_of_schools + resource, data = K12_x) #variables I am curious about + counts
```

```{r, K12 regression}
summary(K12x_5.glm) 
summary(K12x_6.glm) 
summary(K12x_7.glm) #has the smallest AIC and most conclusive results
summary(K12x_8.glm) 

AIC(K12x_5.glm, K12x_6.glm, K12x_7.glm, K12x_8.glm)
BIC(K12x_5.glm, K12x_6.glm, K12x_7.glm, K12x_8.glm)
```

```{r, K12 regression}
#poisson regressionINCONCLUSIVE

#round budget and kWh
K12_x$roundedbudgetperstudent <- round(K12_x$ln_budgetperstudent, 0)
K12_x$roundedkWhperstudent <- round(K12_x$ln_kWhperstudent, 0)

K12x_9.glm <- glm(roundedbudgetperstudent ~ lcff + no_title1 + irr + dac + ces_percentile + ces_score + frmp_count + free_meal_count + obf + income + perc_eligible_free + perc_eligible_frpm + number_of_schools + perc_title1 + enrollment + resource, data = K12_x, family = "poisson") #all variables

K12x_10.glm <- glm(roundedbudgetperstudent ~ lcff + irr + dac + obf + income + perc_eligible_frpm + number_of_schools + perc_title1 + resource, data = K12_x, family = "poisson") #variables I am curious about + perc

K12x_11.glm <- glm(roundedbudgetperstudent ~ no_title1 + frmp_count + free_meal_count +  perc_eligible_frpm + number_of_schools + perc_title1, data = K12_x, family = "poisson") #StepAIC picks 

K12x_12.glm <- glm(roundedbudgetperstudent ~ lcff + no_title1 + irr + dac + frmp_count + obf + income + number_of_schools + resource, data = K12_x, family = "poisson") #variables I am curious about + counts
```

```{r, K12 regression}
#most updated poisson for K12 as of 9/2021
#bc the ln_perstudent values have a right skew, I'm curious to see what poisson gives 

K12.testa <- glm(roundedbudgetper ~ irr + perc_title1, family = "poisson", data = K12_x) #StepAIC variables 

K12.testb <- glm(roundedkWhper ~  lcffperstudent + dac + perc_eligible_frpm, family = "poisson", data = K12_x) #StepAIC variables 

K12.testc <- glm(roundedbudgetper ~ lcffperstudent + irr + dac + perc_eligible_frpm + perc_title1 + obf + resource, family = "poisson", data = K12_x) #intuitive picks based on collinearity 

K12.testd <- glm(roundedkWhper ~ lcffperstudent + irr + dac + perc_eligible_frpm + perc_title1 + obf + resource, family = "poisson", data = K12_x) #intuitive picks based on collinearity 

stargazer(K12.testa, K12.testb, K12.testc, K12.testd, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, K12 regression}
summary(K12x_9.glm) 
summary(K12x_10.glm) 
summary(K12x_11.glm) 
summary(K12x_12.glm) 

AIC(K12x_9.glm, K12x_10.glm, K12x_11.glm, K12x_12.glm)
BIC(K12x_9.glm, K12x_10.glm, K12x_11.glm, K12x_12.glm)
```

```{r, K12 NA}
#poisson budget regression 

K12fu_2.glm <- glm(roundedbudget ~ lcff + no_title1 + irr + dac + frmp_count + obf, data = K12, family = "poisson") 

K12fu_3.glm <- glm(roundedbudget ~ lcff + no_title1 + ces_percentile + ces_score + frmp_count + obf + income + number_of_schools + enrollment + perc_ba + perc_aa, data = K12, family = "poisson") 

K12fu_4.glm <- glm(roundedbudget ~ lcff + no_title1 + irr + dac + frmp_count + obf + income + resource + perc_ia + perc_na + perc_ba + perc_wa + perc_aa, data = K12, family = "poisson") 

summary(K12fu_2.glm) 
summary(K12fu_3.glm) 
summary(K12fu_4.glm)

AIC(K12fu_2.glm, K12fu_3.glm, K12fu_4.glm)
BIC(K12fu_2.glm, K12fu_3.glm, K12fu_4.glm)

```
```{r, model validation !!!}
exp(summary(K12fu_4.glm)$coefficients[,1])
K12$predict_b <- predict(K12fu_4.glm, K12, type = "response")

chisq.test(K12$budget, K12$predict_b) #note, try out with models with limited variables as well 
```

```{r}
plot(K12fu_4.glm)

library('fitdistrplus')
plot(fitdist(K12$roundedbudget,"pois"))
```

```{r, K12 NA}
#negative binomial budget regression 

K12fu_6.glm <- glm(ybudget ~ lcff + no_title1 + irr + dac + frpm_count + obf, data = K12, family = "binomial") 

K12fu_7.glm <- glm(ybudget ~ lcff + no_title1 + ces_percentile + ces_score + frpm_count + obf + income + number_of_schools + enrollment + perc_ba + perc_aa, data = K12, family = "binomial") 

K12fu_8.glm <- glm(ybudget ~ tot_lcff + no_title1_status + irr + dac_proportion + frpm_count + perc_obf + median_household_income_usd + perc_resource + perc_ia + perc_na + perc_ba + perc_wa + perc_aa, data = K12, family = "binomial") #all variables, note that income/ces_score/resource has a high correlation 

summary(K12fu_6.glm) #p-value should be < 0.05 to be significant 
summary(K12fu_7.glm) 
summary(K12fu_8.glm) 

AIC(K12fu_6.glm, K12fu_7.glm, K12fu_8.glm)
AIC(K12fu_6.glm, K12fu_7.glm, K12fu_8.glm)
```

```{r, K12 NA}
#hurdle budget regression
#learn more: https://data.library.virginia.edu/getting-started-with-hurdle-models/

K12fu_10.glm <- hurdle(roundedbudget ~ tot_lcff + no_title1_status + irr + dac_proportion + frpm_count + perc_obf, data = K12) 

K12fu_11.glm <- hurdle(ybudget ~ tot_lcff + no_title1_status + ces_percentile_median + ces_score_median + frpm_count + perc_obf + median_household_income_usd + number_of_schools + enrollment + perc_ba + perc_aa, data = K12) 

K12fu_12.glm <- hurdle(roundedbudget ~ tot_lcff + no_title1_status + irr + dac_proportion + frpm_count + perc_obf + median_household_income_usd + perc_resource + perc_ia + perc_na + perc_ba + perc_wa + perc_aa, data = K12) #all variables, note that income/ces_score/resource has a high correlation 

summary(K12fu_10.glm) #p-value should be < 0.05 to be significant 
summary(K12fu_11.glm) 
summary(K12fu_12.glm)

AIC(K12fu_10.glm, K12fu_11.glm, K12fu_12.glm)
AIC(K12fu_10.glm, K12fu_11.glm, K12fu_12.glm)
```

```{r, K12 NA}
#spline budget regression
#learn more: https://medium.com/analytics-vidhya/spline-regression-in-r-960ca82aa62c

K12fu_14.glm <- lm(budget ~ splines::bs(ln_enrollment, knots = c(8.707481,10.336310, 11.532670)) + tot_lcff + no_title1_status + irr + dac_proportion + frpm_count + perc_obf, data = K12) 

K12fu_15.glm <- lm(budget ~ splines::bs(ln_enrollment, knots = c(8.707481,10.336310, 11.532670)) + tot_lcff + no_title1_status + ces_percentile_median + ces_score_median + frpm_count + perc_obf + median_household_income_usd + number_of_schools + enrollment + perc_ba + perc_aa, data = K12) 

K12fu_16.glm <- lm(budget ~ splines::bs(ln_enrollment, knots = c(8.707481,10.336310, 11.532670)) + tot_lcff + no_title1_status + irr + dac_proportion + frpm_count + perc_obf + median_household_income_usd + perc_resource + perc_ia + perc_na + perc_ba + perc_wa + perc_aa, data = K12) #all variables, note that income/ces_score/resource has a high correlation 

summary(K12fu_14.glm) #p-value should be < 0.05 to be significant 
summary(K12fu_15.glm)
summary(K12fu_16.glm)

AIC(K12fu_14.glm, K12fu_15.glm, K12fu_16.glm)
BIC(K12fu_14.glm, K12fu_15.glm, K12fu_16.glm)
```

```{r, K12 5}

#K12 subset ungrouped

#performs stepwise model selection by AIC
stepAIC(lm(budget ~ tot_lcff + no_title1_status + irr + dac_proportion + ces_percentile_median + ces_score_median + frpm_count + free_meal_count + perc_obf + median_household_income_usd + perc_eligible_free + perc_eligible_frpm + number_of_schools + perc_title1_status + enrollment + perc_resource, data = K12_subset))
```

```{r, this one}
#lcff + no_title1 + irr + dac + ces_percentile + ces_score + frmp_count + free_meal_count + obf + income + perc_eligible_free + perc_eligible_frpm + number_of_schools + perc_title1 + enrollment + resource + perc_ia + perc_na + perc_ba + perc_wa + perc_aa

K12s_5.glm <- glm(budgetperstudent ~ lcff + no_title1 + frmp_count + income + irr + dac + obf + resource, data = K12_x) 
summary(K12s_5.glm)

exp(summary(K12s_5.glm)$coefficients[,1])
K12_subset$predict_b <- predict(K12s_5.glm, K12_subset, type = "response")

chisq.test(K12_subset$budget, K12_subset$predict_b)
```

```{r, this one}
K12s_6.glm <- glm(kWhperstudent ~ lcffperstudent + perc_title1 + perc_eligible_frpm + income + irr + dac + obf + resource, data = K12_x) 
summary(K12s_6.glm)

exp(summary(K12s_6.glm)$coefficients[,1])
K12_subset$predict_b <- predict(K12s_6.glm, K12_subset, type = "response")

chisq.test(K12_subset$budget, K12_subset$predict_b)
```

```{r, this one}
#t-tests
t.test(K12_no$irr, K12_subset$irr, na.action = na.omit)
#t.test(no_participate$budgetperstudent, participate$budgetperstudent, na.action = na.omit)
#t.test(no_participate$kWhperstudent, participate$kWhperstudent, na.action = na.omit)
```

```{r, K12 }
#linear budget regression 

K12s_2.glm <- glm(lnbudgetperschool ~ lcffperschool + no_title1_status + ces_percentile_median + ces_score_median + frpm_count + perc_eligible_free + perc_eligible_frpm + perc_ba + perc_aa, data = K12_x) #StepAIC

K12s_3.glm <- glm(ln_budget ~ tot_lcff + no_title1_status + irr + dac_proportion + perc_eligible_frpm + perc_obf, data = K12_x) #variables I am curious about 

K12s_4.glm <- glm(lnbudgetperschool ~ lcffperschool + no_title1_status + irr + dac_proportion + frpm_count + perc_obf + median_household_income_usd + perc_resource + perc_ia + perc_na + perc_ba + perc_wa + perc_aa, data = K12_x) #all variables, note that income/ces_score/resource has a high correlation 

summary(K12s_2.glm) #p-value should be < 0.05 to be significant 
summary(K12s_3.glm) 
summary(K12s_4.glm) 

AIC(K12s_2.glm, K12s_3.glm, K12s_4.glm)
BIC(K12s_2.glm, K12s_3.glm, K12s_4.glm)
```

```{r}
#Checking to see if the regression actually fits the data
with(counties1a_glm.nb, cbind(res.deviance = deviance, df = df.residual,
  p = pchisq(deviance, df.residual, lower.tail=TRUE)))
with(counties1b_glm.nb, cbind(res.deviance = deviance, df = df.residual,
  p = pchisq(deviance, df.residual, lower.tail=TRUE)))
with(counties1c_glm.nb, cbind(res.deviance = deviance, df = df.residual,
  p = pchisq(deviance, df.residual, lower.tail=TRUE)))
with(counties1d_glm.nb, cbind(res.deviance = deviance, df = df.residual,
  p = pchisq(deviance, df.residual, lower.tail=TRUE)))
with(counties1e_glm.nb, cbind(res.deviance = deviance, df = df.residual,
  p = pchisq(deviance, df.residual, lower.tail=TRUE)))
```

```{r, counties regression}
#most updated poisson for counties 9/2021
#bc the percapita values have a right skew, I'm curious to see what poisson gives 

counties.testa <- glm(roundedbudgetper ~ population, family = "poisson", data = counties_x) #StepAIC variables 

counties.testb <- glm(roundedkWhper ~  revpercapita + population, family = "poisson", data = counties_x) #StepAIC variables 

counties.testc <- glm(roundedbudgetper ~ revpercapita + dac + irr + obf + income, family = "poisson", data = counties_x) #intuitive picks based on collinearity 

counties.testd <- glm(roundedkWhper ~ revpercapita + dac + irr + obf + income, family = "poisson", data = counties_x) #intuitive picks based on collinearity 

stargazer(counties.testa, counties.testb, counties.testc, counties.testd, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```

```{r, cities regression}
#most updated poisson for counties 9/2021
#bc the percapita values have a right skew, I'm curious to see what poisson gives 

cities.testa <- glm(roundedbudgetper ~ revpercapita + ruca + dac + obf +  income + resource, family = "poisson", data = cities_x) #StepAIC variables 

cities.testb <- glm(roundedkWhper ~  revpercapita + ruca + dac + obf + income, family = "poisson", data = cities_x) #StepAIC variables 

cities.testc <- glm(roundedbudgetper ~ revpercapita + ruca + dac + population + obf + income + resource, family = "poisson", data = cities_x) #intuitive picks based on collinearity 

cities.testd <- glm(roundedkWhper ~ revpercapita + ruca + dac + population + obf + income + resource , family = "poisson", data = cities_x) #intuitive picks based on collinearity 

stargazer(cities.testa, cities.testb, cities.testc, cities.testd, type="text", title="Results", align=TRUE, omit.stat=c("LL","ser","f"), no.space=TRUE)
```
